Using tours to visually investigate properties of new projection pursuit indexes with application to problems in physics

### 2018-11-21 scagnostics.Rmd
tidied data section

The structured projection is in fifth and sixth variables. Two samples sizes are also used: $n=(100, 1000)$. All variables are standardized to mean 0 and standard deviation 1.

@@

Figure \ref{data} shows samples from each of the families, of the nuisance and structured pairs of variables.

@@

@@

Table \ref{tab:indexTable} compares the PPIs for structured projections against those for nuisance variables, based on 100 simulated data sets of each type, using sample size 1000. The lower and upper show the 5th and 95th percentile of values. All indexes, except convex show distinctly higher values for the structured projections. The convex index considers the noise projections more interesting than the structured projections, which might be expected, but also provides an argument to use (1-convex) as the PPI.

@@

\textcolor{red}{XXX This shouldn't go here. The index needs to have been introduced much earlier in the paper. TIC was suggested in @JMLRv1715308 as an alternative to MIC with better properties for testing against independence, this is because the full characteristic matrix is used, implying that more information is considered. We also see that TIC is smoother when moving into the special view, in terms of the relative differences.}
<!--
We repeat the same exercise, but now we are interested in how the index values change when going from the view in $x_1$ vs $x_2$ to the view in $x_5$ vs $x_6$, which will also allow us to view the evolution between typical low and high values of the index function. Results are shown in Figure \ref{fig:plannedtour}.
We can see a pretty clear and smooth increase in dcor2d and splines2d as well as MIC and TIC index when approaching the special view of the Sine distribution, which should therefore be efficient to optimise. On the other hand while we also see smooth increase in MIC and TIC when approaching the special view of the Pipe distribution, the increase only becomes apparent much closer to the final projection, making this structure more challenging to detect with a guided tour. Finally the Spiral appears to be the most challenging, having an even smaller squint angle than the Pipe distribution. While "cmass" appears to be appropriate to identify this view we may actually find larger values of this index for random distributions, making it unfit to use in a guided tour. The scagnostics indices skinny (or 1-convex) appear to be sensitive to the structure at a larger angle compared to MINE indices, but will suffer from the large fluctuations in random views when used in a guided tour.
Various indices appear less noisy than seen in \ref{fig:plotV1V2toV3V4}, but note that the y-axis scale is different (typically reaching higher index values in this example, whereas smaller fluctuations are more amplified in the previous views).

@@

## Effect of parameter choice in nndex value
\label{sec:params}
Some parameters must be fixed when using the index functions considered here, controlling trade-offs between noise and fine structure detection, and between computing time and precision.

>>>>

The structured projection is in variables 5 and 6. Two samples sizes are used: $n=(100, 1000)$. All variables are standardized to have mean 0 and standard deviation 1.

@@

@@

Figure \ref{fig:data} shows samples from each of the families, of the nuisance and structured pairs of variables. Table \ref{tab:indexTable} compares the PPIs for structured projections against those for nuisance variables, based on 100 simulated data sets of each type, using sample size 1000. The lower and upper show the 5th and 95th percentile of values. All indexes, except convex show distinctly higher values for the structured projections. The convex index considers the noise projections more interesting than the structured projections, which might be expected, but also provides an argument to use (1-convex) as the PPI.

@@

@@

<!--
\textcolor{red}{XXX This shouldn't go here. The index needs to have been introduced much earlier in the paper. TIC was suggested in @JMLRv1715308 as an alternative to MIC with better properties for testing against independence, this is because the full characteristic matrix is used, implying that more information is considered. We also see that TIC is smoother when moving into the special view, in terms of the relative differences.}
-->

<!--
We repeat the same exercise, but now we are interested in how the index values change when going from the view in $x_1$ vs $x_2$ to the view in $x_5$ vs $x_6$, which will also allow us to view the evolution between typical low and high values of the index function. Results are shown in Figure \ref{fig:plannedtour}.
We can see a pretty clear and smooth increase in dcor2d and splines2d as well as MIC and TIC index when approaching the special view of the Sine distribution, which should therefore be efficient to optimise. On the other hand while we also see smooth increase in MIC and TIC when approaching the special view of the Pipe distribution, the increase only becomes apparent much closer to the final projection, making this structure more challenging to detect with a guided tour. Finally the Spiral appears to be the most challenging, having an even smaller squint angle than the Pipe distribution. While "cmass" appears to be appropriate to identify this view we may actually find larger values of this index for random distributions, making it unfit to use in a guided tour. The scagnostics indices skinny (or 1-convex) appear to be sensitive to the structure at a larger angle compared to MINE indices, but will suffer from the large fluctuations in random views when used in a guided tour.
Various indices appear less noisy than seen in \ref{fig:plotV1V2toV3V4}, but note that the y-axis scale is different (typically reaching higher index values in this example, whereas smaller fluctuations are more amplified in the previous views).

@@

## Effect of parameter choice in index value
\label{sec:params}
Some parameters must be fixed when using the index functions considered here, controlling trade-offs between noise and fine structure detection, and between computing time and precision.
